---
title: "Main Functions in tidyplastic"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Main Functions in tidyplastic}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Overview
The `tidyplastic` package provides tools for analyzing plastic 
waste cleanup efficiency using the Break Free From Plastic dataset.
Below are the main functions with examples of usage.

## Function 1: `load_data()`
Loads the Break Free From Plastic dataset from the bundled CSV file
in the package. This is the starting point for all other functions.

```{r}
#| eval: false
plastics <- load_data()
```

## Function 2: `clean_plastic_data()`
Standardizes country names and joins region and population data
from the REST Countries API. This is the data cleaning step required
before computing efficiency or summarizing by region.

```{r}
#| eval: false
plastics <- load_data()
cleaned <- clean_plastic_data(plastics)
```

## Function 3: `compute_cleanup_efficiency()`
A summary calculation that aggregates the raw plastics dataset to the
country-year level and calculates cleanup efficiency as total plastic
collected per volunteer. This condenses thousands of brand-level rows
into one efficiency metric per country.

```{r}
#| eval: false
plastics <- load_data()
cleaned <- clean_plastic_data(plastics)
efficiency <- compute_cleanup_efficiency(cleaned)
```

## Function 4: `summarize_by_region()`
Aggregates plastic waste and cleanup event data to the region level,
computing total plastic collected, total events, and average cleanup
efficiency per region. Useful for comparing cleanup performance across
different parts of the world.

```{r}
#| eval: false
plastics <- load_data()
cleaned <- clean_plastic_data(plastics)
summarize_by_region(cleaned)
```

## Function 5: `join_gdp()`
Enriches a cleanup efficiency tibble with GDP per capita data from
the bundled plotting data file. Adds `iso` and `gdp_per_capita_nominal`
columns to the efficiency tibble.

```{r}
#| eval: false
plastics <- load_data()
cleaned <- clean_plastic_data(plastics)
efficiency <- compute_cleanup_efficiency(cleaned)
enriched <- join_gdp(efficiency)
```

## Function 6: `plot_cleanup_efficiency()`
A visualization presenting a linear model with density bins instead
of points. Depicts the relationship between cleanup efficiency and
volunteers on log-log axes. Useful for answering: does increasing
the number of volunteers correlate with an increase in cleanup efficiency?

```{r}
#| eval: false
plastics <- load_data()
cleaned <- clean_plastic_data(plastics)
efficiency <- compute_cleanup_efficiency(cleaned)
plot_cleanup_efficiency(efficiency)
```

## Function 7: `plot_gdp_efficiency()`
An interactive scatter plot of GDP per capita vs cleanup efficiency,
colored by region. Useful for identifying regional patterns and
outliers like Togo.

```{r}
#| eval: false
plastics <- load_data()
cleaned <- clean_plastic_data(plastics)
efficiency <- compute_cleanup_efficiency(cleaned)
enriched <- join_gdp(efficiency)
plot_gdp_efficiency(enriched)
```

## Function 8: `test_gdp_plastic_association()`
Runs a chi-square test to assess whether GDP group and dominant plastic
type are statistically associated. Returns a tibble with the chi-square
statistic, degrees of freedom, and p-value.

```{r}
#| eval: false
plastics <- load_data()
cleaned <- clean_plastic_data(plastics)
efficiency <- compute_cleanup_efficiency(cleaned)
enriched <- join_gdp(efficiency)
test_gdp_plastic_association(enriched)
```
