---
title: "Main Functions in tidyplastic"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Main Functions in tidyplastic}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
## Overview

The `tidyplastic` package provides tools for analyzing plastic 
waste cleanup efficiency using the Break Free From Plastic dataset.
Below are three planned functions we have with examples of aspirational usage.

## Function 1: `compute_cleanup_efficiency()`

A summary calculation that aggregates the raw plastics dataset to the
country-year level and calculates cleanup efficiency as total plastic
collected per volunteer. This condenses thousands of brand-level rows
into one efficiency metric per country.

```{r}
#| eval: false
plastics <- load_data()
compute_cleanup_efficiency(plastics)
```

## Function 2: `join_gdp()`

Makes a cleanup efficiency tibble with GDP per capita data using
the API Ninjas GDP endpoint. Falls back to a proxy score based on
volunteer and plastic totals if the API is unavailable. The gdp_source
column tracks whether each value came from the API or the fallback proxy.

```{r}
#| eval: false
plastics <- load_data()
efficiency <- compute_cleanup_efficiency(plastics)
join_gdp(efficiency, year = 2019)
```

## Function 3: `plot_gdp_efficiency()`

A visualization that produces an interactive scatter plot of GDP per
capita vs cleanup efficiency, colored by region, with an optional
per-region linear line of best fit. This would be useful for identifying regional patterns and outliers like Togo.

```{r}
#| eval: false
plastics <- load_data()
efficiency <- compute_cleanup_efficiency(plastics)
enriched <- join_gdp(efficiency, year = 2019)
plot_gdp_efficiency(enriched, regions = c("africa", "asia", "europe"),
                    log_axes = TRUE, show_smoother = TRUE)
```
